Understanding Factors Contributing to Application Approval in Urban Planning Decisions (2016-2024)

### 2024-12-03 paper/paper.qmd
Upload datasheet

---
title: "Understanding Factors Contributing to Application Approval in Urban Planning Decisions (2016-2024)"
subtitle: "Analysis using data from Toronto’s Committee of Adjustment applications using Bayesian Logistic Regression"
author: Angel Xu
thanks: "Code and data are available at: [https://github.com/Anjojoo/Committee-of-Adjustment-Applications](https://github.com/Anjojoo/Committee-of-Adjustment-Applications)."
date: today
date-format: long
abstract: "This paper analyzes urban planning application approvals submitted to Toronto's Committee of Adjustment from 2016 to 2024, focusing on key predictors such as application type, submission year, and planning district using Bayesian logistic regression. The analysis shows that consents are more likely to be approved than minor variances, with approval rates generally increasing over time, particularly from 2017 onward, though fluctuations are observed in certain years. Geographic disparities were identified, with North York showing the highest likelihood of approval, while Scarborough faced stricter reviews. These findings enhance understanding of urban planning decision-making processes, offering insights into temporal trends, geographic variability, and areas for equitable policy adjustments."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

#### Plot the distribution of `decision` ####
ggplot(analysis_data, aes(x = factor(decision))) +
  geom_bar(fill = "lightblue", color = "black") +
  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5) +
  labs(
    x = "Decision (0 = Refused, 1 = Approved)",
    y = "Count"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    axis.text = element_text(size = 12)
  )

```

@@

#### Generate summary statistics for `decision` ####
decision_summary <- analysis_data %>%
  group_by(decision) %>%
  summarize(
    Percentage = paste0(round((n() / nrow(analysis_data)) * 100, 1), "%") # Add '%' symbol
  ) %>%
  mutate(Decision = ifelse(decision == 1, "Approved (1)", "Refused (0)")) %>%
  select(Decision, Percentage)

@@

#### Plot the distribution of `application_type` ####
ggplot(analysis_data, aes(x = factor(application_type))) +
  geom_bar(fill = "lightgreen", color = "black") +
  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5) +
  labs(
    x = "Application Type",
    y = "Count"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    axis.text = element_text(size = 12)
  )

```

@@

#### Summary Statistics for Application Type ####
application_summary <- analysis_data %>%
  group_by(application_type) %>%
  summarize(
    Percentage = paste0(round((n() / nrow(analysis_data)) * 100, 1), "%") # Add %
  ) %>%
  mutate(Application_Type = ifelse(application_type == "MV", "Minor Variance (MV)", "Consent (CO)")) %>%
  select(Application_Type, Percentage)

@@

#### Plot the distribution of `year` ####
ggplot(analysis_data, aes(x = factor(year))) +
  geom_bar(fill = "skyblue", color = "black") +
  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5) +
  labs(
    x = "Year",
    y = "Count"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    axis.text = element_text(size = 12)
  )

```

@@

#### Plot the distribution of `planning_district` ####
ggplot(analysis_data, aes(x = factor(planning_district))) +
  geom_bar(fill = "coral", color = "black") +
  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5) +
  labs(
    x = "Planning District",
    y = "Count"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    axis.text = element_text(size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1) # Rotate district labels for readability
  )

```

@@

@@

# Plot boxplot for application type
ggplot(predict_data_long, aes(x = application_type, y = prob)) +
  geom_boxplot() +
  labs(
    x = "Application Type",
    y = "Predicted Probability of Approval"
  ) +
  theme_minimal()

```

@@

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 1
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

@@

This method reduces sampling variability, enhances representation, and provides more accurate estimates compared to simple random sampling, as the unique characteristics of each subgroup are reflected in the final sample.

@@

- Your answers will be treated with complete confidentiality.
- Participation in this survey is voluntary.
- You are encouraged to provide honest and thoughtful responses.
- The survey is estimated to take about 5 minutes to complete.
- If you have any questions or concerns, feel free to reach out to me at [anjojoo.xu@mail.utoronto.ca](anjojoo.xu@mail.utoronto.ca) (Angel Xu).

@@

If you have any questions or would like to learn more about this study, please don’t hesitate to contact our research team at [anjojoo.xu@mail.utoronto.ca](anjojoo.xu@mail.utoronto.ca) (Angel Xu).

>>>>

---
title: "Understanding Factors Contributing to Application Approval in Urban Planning Decisions (2016-2024)"
subtitle: "Analysis using data from Toronto’s Committee of Adjustment applications using Bayesian Logistic Regression"
author: Victoria Zheng
thanks: "Code and data are available at: [https://github.com/Anjojoo/Committee-of-Adjustment-Applications](https://github.com/Anjojoo/Committee-of-Adjustment-Applications)."
date: today
date-format: long
abstract: "This paper analyzes urban planning application approvals submitted to Toronto's Committee of Adjustment from 2016 to 2024, focusing on key predictors such as application type, submission year, and planning district using Bayesian logistic regression. The analysis shows that consents are more likely to be approved than minor variances, with approval rates generally increasing over time, particularly from 2017 onward, though fluctuations are observed in certain years. Geographic disparities were identified, with North York showing the highest likelihood of approval, while Scarborough faced stricter reviews. These findings enhance understanding of urban planning decision-making processes, offering insights into temporal trends, geographic variability, and areas for equitable policy adjustments."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

#### Plot the distribution of `decision` ####
ggplot(analysis_data, aes(x = factor(decision))) +
  geom_bar(fill = "lightblue", color = "black") +
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5) +
  labs(
    x = "Decision (0 = Refused, 1 = Approved)",
    y = "Count"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    axis.text = element_text(size = 12)
  )
```

@@

#### Generate summary statistics for `decision` ####
decision_summary <- analysis_data %>%
  group_by(decision) %>%
  summarize(
    Percentage = paste0(round((n() / nrow(analysis_data)) * 100, 1), "%")
  ) %>%
  mutate(Decision = ifelse(decision == 1, "Approved (1)", "Refused (0)")) %>%
  select(Decision, Percentage)

@@

#### Plot the distribution of `application_type` ####
ggplot(analysis_data, aes(x = factor(application_type))) +
  geom_bar(fill = "lightgreen", color = "black") +
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5) +
  labs(
    x = "Application Type",
    y = "Count"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    axis.text = element_text(size = 12)
  )
```

@@

#### Summary Statistics for Application Type ####
application_summary <- analysis_data %>%
  group_by(application_type) %>%
  summarize(
    Percentage = paste0(round((n() / nrow(analysis_data)) * 100, 1), "%")
  ) %>%
  mutate(Application_Type = ifelse(application_type == "MV",
    "Minor Variance (MV)", "Consent (CO)"
  )) %>%
  select(Application_Type, Percentage)

@@

#### Plot the distribution of `year` ####
ggplot(analysis_data, aes(x = factor(year))) +
  geom_bar(fill = "skyblue", color = "black") +
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5) +
  labs(
    x = "Year",
    y = "Count"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    axis.text = element_text(size = 12)
  )
```

@@

#### Plot the distribution of `planning_district` ####
ggplot(analysis_data, aes(x = factor(planning_district))) +
  geom_bar(fill = "coral", color = "black") +
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5) +
  labs(
    x = "Planning District",
    y = "Count"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    axis.text = element_text(size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

@@

`Mostly Harmless Econometrics` (@angrist2009mostly) addresses how to handle variability in time-series and cross-sectional data, focusing on ensuring robust and interpretable results. The use of submission year and planning district as key predictors aligns with Angrist and Pischke's guidance on analyzing temporal and geographic variation. Their discussion on the importance of controlling for time trends and location-specific effects validates the inclusion of these variables in the Bayesian logistic regression model.

@@

# Plot boxplot for application type
ggplot(predict_data_long, aes(x = application_type, y = prob)) +
  geom_boxplot() +
  labs(
    x = "Application Type",
    y = "Predicted Probability of Approval"
  ) +
  theme_minimal()
```

@@

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 1
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check",
#| "Comparing the posterior with the prior"]

@@

`Sampling: Design and Analysis` (@lohr2021sampling) explores how stratified sampling reduces variability by dividing populations into homogeneous groups, enabling more precise estimates and better resource allocation. By applying these principles, the paper ensures that the analysis accounts for heterogeneity in geographic and temporal patterns of planning applications, enabling robust conclusions about decision-making trends. The methodology follows Lohr’s recommendation to tailor strata definitions to the study's objectives, enhancing the reliability of the results.

Besides, `Sampling` (@thompson2012sampling) provides a detailed guide to modern sampling techniques, emphasizing the importance of tailoring approaches to study objectives. His discussion on stratified sampling highlights its benefits in improving precision and reducing bias by dividing populations into homogeneous subgroups. This aligns with the choice of using planning districts, application types, and submission years as strata to ensure balanced representation and capture temporal and spatial variability in urban planning decisions. Thompson’s principles validate the proportional sampling approach adopted here, supporting its effectiveness in generating reliable analyses.

@@

- Your answers will be treated with complete confidentiality.
- Participation in this survey is voluntary.
- You are encouraged to provide honest and thoughtful responses.
- The survey is estimated to take about 5 minutes to complete.
- If you have any questions or concerns, feel free to reach out to me at [anjojoo.xu@mail.utoronto.ca](anjojoo.xu@mail.utoronto.ca) (Victoria Zheng).

@@

If you have any questions or would like to learn more about this study, please don’t hesitate to contact our research team at [anjojoo.xu@mail.utoronto.ca](anjojoo.xu@mail.utoronto.ca) (Victoria Zheng).
